Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-25 src/5.Example.tex
Bug fixes.

In this section we provide an example of an application written in the Kokkos programming model. The code in Figure~\ref{fig:KokkosExample} shows a parallel matrix-vector multiplication implemented as two nested loops. The outer loop, that is the outer parallel_for, iterates over the rows of the input matrix and of the input vector while the inner loop iterates over elements in each row. Since the inner loop potentially executes in parallel, the update operation of the aggregating variable~\emph{y_tmp} must be protected against data races. Using the parallel_reduction pattern, the programming model provides a local variable to that loop iteration. Once the computation finishes, the final result of the reduction operation is stored in the output vector~\emph{y} at the row index~\emph{e}.  
As a first step, a Kokkos application must call ~\emph{Kokkos::initialize}. Initialization and finalization are required and serve the purpose of setting up the underlying runtime. 
Further, the example code shows the use of the ~\emph{Kokkos::View} abstraction.

@@

Taking a closer look at the loop step shows the hierarchical expression of concurrency. In case of the matrix-vector multiplication,This is accomplished by using the ~\emph{Kokkos::TeamPolicy}. For each row of~\emph{M} a team of threads is launched, where the size of the team is not further specified. When using a ~\emph{TeamPolicy} the operator of the lambda expression does not receive an index, but rather a handle to the team of threads. 
This handle provides the team identifier \emph{e}, and is subsequently handed to the nested parallel execution policies. The nested reduction requires the operator of the nested lambda to take a reference to the thread-local reduction variable (called here ~\emph{y\_tmp}) in addition to the loop index. It then writes the result to the correct position of the output vector. 
The parallel loops are follows by a call to \emph{Kokkos::fence}. This ensures that the execution of the parallel code is completed before copying the results back to the host memory space. This is implemented through the~\emph{Kokkos::deep_copy} call. Lastly, the Kokkos application calls~\emph{Kokkos::finalize}.

@@

In this section we peak inside the Kokkos library and show how the a generic parallel for loop is mapped to the OpenMP back-end programming model. As a template metaprogramming library, Kokkos makes use of partial specialization. In particular the~\emph{parallel_for} or ~\emph{parallel_reduce} functional calls as shown in Figure~\ref{fig:KokkosExample} are mapped to partially specialized classes. Such classes correspond to particular back-end where for each back-end, concepts with partial specialization must be provided. Today, Kokkos supports the CUDA and OpenMP back-ends. For each new back-end, partially implemented classes that implement necessary concepts must be provided.

>>>>

In this section we provide an example of an application written in the Kokkos programming model. The code in Figure~\ref{fig:KokkosExample} shows a parallel matrix-vector multiplication implemented as two nested loops. The outer loop, that is the outer~\emph{parallel\_for}, iterates over the rows of the input matrix and of the input vector while the inner loop iterates over elements in each row. Since the inner loop potentially executes in parallel, the update operation of the aggregating variable~\emph{y\_tmp} must be protected against data races. Using the~\emph{parallel\_reduction} pattern, the programming model provides a local variable to that loop iteration. Once the computation finishes, the final result of the reduction operation is stored in the output vector~\emph{y} at the row index~\emph{e}.  
As a first step, a Kokkos application must call ~\emph{Kokkos::initialize}. Initialization and finalization are required and serve the purpose of setting up the underlying runtime. 
Further, the example code shows the use of the ~\emph{Kokkos::View} abstraction.

@@

Taking a closer look at the loop step shows the hierarchical expression of concurrency. In case of the matrix-vector multiplication,This is accomplished by using the ~\emph{Kokkos::TeamPolicy}. For each row of~\emph{M} a team of threads is launched, where the size of the team is not further specified. When using a ~\emph{TeamPolicy} the operator of the lambda expression does not receive an index, but rather a handle to the team of threads. 
This handle provides the team identifier \emph{e}, and is subsequently handed to the nested parallel execution policies. The nested reduction requires the operator of the nested lambda to take a reference to the thread-local reduction variable (called here ~\emph{y\_tmp}) in addition to the loop index. It then writes the result to the correct position of the output vector. 
The parallel loops are follows by a call to \emph{Kokkos::fence}. This ensures that the execution of the parallel code is completed before copying the results back to the host memory space. This is implemented through the~\emph{Kokkos::deep\_copy} call. Lastly, the Kokkos application calls~\emph{Kokkos::finalize}.

@@

In this section we peak inside the Kokkos library and show how the a generic parallel for loop is mapped to the OpenMP back-end programming model. As a template metaprogramming library, Kokkos makes use of partial specialization. In particular the~\emph{parallel\_for} or ~\emph{parallel\_reduce} functional calls as shown in Figure~\ref{fig:KokkosExample} are mapped to partially specialized classes. Such classes correspond to particular back-end where for each back-end, concepts with partial specialization must be provided. Today, Kokkos supports the CUDA and OpenMP back-ends. For each new back-end, partially implemented classes that implement necessary concepts must be provided.
